---
title: "HW2"
author: "Alexandra Gibbons"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    theme: united
    keep_md: TRUE
   
---
```{r, setup, include=FALSE}
# MIGHT NOT NEED ANYMORE
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/ajg109//Documents/Github/223_course")
knitr::opts_chunk$set(message = FALSE)
```

# MD 3

```{r}
library(tidyverse)
mario_kart <- read_csv("~/Documents/GitHub/223_course/Data/world_records.csv")
glimpse(mario_kart)
```

## Question 1
```{r}
three_laps <- mario_kart %>% filter(type == "Three Lap")
no_rainbow_rd <- three_laps %>% filter(track != "Rainbow Road")
rainbow_rd <- three_laps %>%  filter(track == "Rainbow Road")
```

Above, I use the filter function to create one dataset containing only Rainbow Road records, and one dataset that excludes all Rainbow Road records. 

## Question 2
```{r}
rainbow_rd_summary <- rainbow_rd %>% summarise (rainbow_mean=mean(time), rainbow_sd=sd(time))
rainbow_rd_summary
all_others_summary <- no_rainbow_rd %>% summarise(all_others_mean=mean(time), all_others_sd=sd(time))
all_others_summary
```

The Rainbow Road records have a longer mean time and larger standard deviation compared to the mean and standard deviation completion time for all other tracks combined.

## Question 3
```{r}
track_records <- three_laps %>% 
  group_by(track) %>% 
  summarize(record_count=n()) %>% 
  arrange(desc(record_count))
track_records
```

Toad's Turnpike has the most records established in it.

## Question 4
```{r}
driver_records <- three_laps %>% 
  group_by(player, track) %>% 
  summarize(record_count=n()) %>% 
  arrange(desc(record_count))
driver_records
```

Penev has recorded the most records at any one track, and that track was Choco Mountain.

## Question 5
```{r}
three_laps %>% 
  group_by(track) %>% 
  summarize(mean_time=mean(time)) %>% 
  arrange(desc(mean_time))
```

The average time for each track is shown above. Rainbow Road, as expected, has the longest average time.

```{r}
ungroup(three_laps)
three_laps %>% 
  group_by(track) %>% 
  arrange(time) %>% 
  select(track, time) %>% 
  slice(1)
```

The fastest time for each track is shown above.

## Question 6
```{r}
three_laps %>% 
  mutate(duration_over_100=case_when(record_duration>100 ~ 1,
                                     record_duration<=100 ~ 0)) %>% 
  group_by(player) %>% 
  summarize(long_dur_count=sum(duration_over_100)) %>% 
  arrange(desc(long_dur_count))
```

The total amount of long duration records (over 100 days) that each player holds are above. MR has the most long duration records (81).

## Question 7
```{r}
drivers <- read_csv("~/Documents/GitHub/223_course/Data/drivers.csv")
mario_joined <- left_join(three_laps, drivers, by='player')
mario_joined %>% 
  count(nation) %>% 
  ggplot(mapping=aes(x=fct_reorder(nation, n), y=n)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x="Nation", y="Number of Mario Kart Records", 
       title="Number of Mario Kart Records by Players' Nation") 
```

# MD 4

## Question 1
```{r}
nfl_salaries <- read_csv("https://raw.githubusercontent.com/NicolasRestrep/223_course/main/Data/nfl_salaries.csv")
nfl_salaries
```

## Question 2
```{r}
nfl_salaries_tidy <- nfl_salaries %>% 
  pivot_longer(names_to= "position",
               values_to= "salary",
               cols= -year)
nfl_salaries_tidy
```
## Question 3
```{r}
quarterbacks <- nfl_salaries_tidy %>% 
  filter(position=="Quarterback")

ggplot(data=quarterbacks, mapping=aes(x=salary)) +
  geom_histogram(color="white", binwidth = 1000000) + 
  scale_x_continuous(name="NFL Quarterback Salaries", labels = scales::comma)

ggplot(data=quarterbacks, mapping=aes(x=salary)) +
  geom_histogram(color="white", binwidth = 1000000) + 
  scale_x_continuous(name="NFL Quarterback Salaries", labels = scales::comma) +
  facet_wrap(~year) +
  theme(axis.text.x = element_text(size = 7)) 
```

Over time, there has been an increase in extremely high NFL Quarterback salaries. In 2011, there were not salaries above 20 million, yet we see the prevalence of these high salaries steadily increase over time. 

## Question 4
```{r}
avg_nfl_salaries <- nfl_salaries_tidy %>% 
  group_by(position, year) %>% 
  summarise(avg_salary=mean(salary, na.rm=TRUE))
avg_nfl_salaries
```

## Question 5
```{r}
library("RColorBrewer")
ggplot(data=avg_nfl_salaries, mapping=aes(x=year, y=avg_salary, group=position, color=position)) +
         geom_line() +
        scale_color_brewer(palette= "Paired") +
  labs(y="average salary", 
       title="Trends in Average NFL Salary by Position, 2011-2018") +
  scale_y_continuous(labels = scales::comma)
```

For most positions, average salaries have increased between 2011 and 2018, with the exception of running backs. The average salaries of offensive linemen and defensive linemen have increased rapidly since 2016. 